Comparing Generative Pre-trained Transformer Models to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-02-13 manuscript/manuscript.tex
Move all vs agree reference from methods to results
Edit names

\author*[1,2]{\fnm{Richard} \sur{Wen}}\email{richard.wen@unityhealth.to}

\author[1,5]{\fnm{Rajeev} \sur{Kamadod}}\email{rajeevk@kentropy.com}

@@

@@

\author[1,3]{\fnm{Asha} \sur{Behdinan}}\email{asha.behdinan@mail.utoronto.ca}

@@

\author[1,3]{\fnm{Areeba} \sur{Zubair}}\email{areeba.zubair@mail.utoronto.ca}

\author[1]{\fnm{Thomas Kai Sze} \sur{Ng}}\email{KaiSze.Ng@unityhealth.to}

\author[1]{\fnm{Andy} \sur{Lee}}\email{Andy.Lee2@unityhealth.to}

\author[1,2]{\fnm{Paijani} \sur{Sheth}}\email{paijani.sheth@mail.utoronto.ca}

\author[1]{\fnm{Anteneh Tesfaye} \sur{Assalif}}\email{antenehta@gmail.com}

\author[1,4]{\fnm{Patrick} \sur{Brown}}\email{}

\author[1,2]{\fnm{Prabhat} \sur{Jha}}\email{prabhat.jha@utoronto.ca}

\affil*[1]{\orgdiv{Centre for Global Health Research, St. Michael's Hospital}, \orgname{Unity Health Toronto}, \orgaddress{\street{30 Bond St}, \city{Toronto}, \postcode{M5B 1W8}, \state{Ontario}, \country{Canada}}}

\affil[2]{\orgdiv{Dalla Lana School of Public Health}, \orgname{University of Toronto}, \orgaddress{\street{155 College Street}, \city{Toronto}, \postcode{M5T 3M7}, \state{Ontario}, \country{Canada}}}

\affil[3]{\orgdiv{Department of Surgery}, \orgname{University of Toronto}, \orgaddress{\street{149 College Street}, \city{Toronto}, \postcode{M5T 1P5}, \state{Ontario}, \country{Canada}}}

\affil[4]{\orgdiv{Department of Statistical Sciences}, \orgname{University of Toronto}, \orgaddress{\street{700 University Avenue}, \city{Toronto}, \postcode{M5G 1Z5}, \state{Ontario}, \country{Canada}}}

\affil[5]{\orgname{Kentropy Technologies Pvt. Ltd.}, \orgaddress{\street{2nd Main Rd}, \city{Bengaluru}, \postcode{560 034}, \state{Bangalore}, \country{India}}}

@@

The CSMF accuracy and PCCC metrics were calculated and compared for each model overall and by age group, followed by age ranges and COD for each age group as model performance can vary across ages and specific causes \cite{benaraEvaluationMethodsAssigning2023,murrayUsingVerbalAutopsy2014,setelValidityVerbalAutopsy2006}. Metrics were calculated overall for three age groups according to the CGHR-10 codes: adult (12 to 69 years), child (28 days to 11 years), and neonatal (under 28 days). For each of the adult and child age groups, metrics were calculated for five-year age ranges for records with ages at death of one-year or older and five-month age ranges for 28 days or older. For the neonatal age group, the age ranges of 0-6 days and 7-27 days were used. Metrics were also calculated by CODs defined by CGHR-10 codes, which include 19, 10, and 7 CODs for adult, child, and neonatal records respectively. For reference, all records were also coded by GPT-3.5, GPT-4, InterVA-5, and InSilicoVA for comparison to physician agreed records. For physician agreed records, models had 0.05 to 0.1 higher PCCC values and 0.01 to 0.05 higher CSMF accuracy values (see Figure \ref{fig-perf-allvsagree} in Appendix \ref{app-perf-misc}).

@@

Of the 6939 records, GPT-4 had the highest PCCC followed by GPT-3.5, InSilicoVA, and InterVA-5 at 0.71, 0.64, 0.44, and 0.43 PCCC respectively (Table \ref{tab-perf-overall}). GPT-3.5 and GPT-4 had improvements ranging from 0.20-0.28 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.07 PCCC. CSMF accuracies were above 0.7 for all models at 0.89, 0.86, 0.74, and 0.78 for GPT-4, GPT-3.5, InSilicoVA, and InterVA-5 respectively. Figure \ref{fig-perf-agegroup} shows the PCCC performance across age groups with reference to the CSMF Accuracy. GPT-4 had the highest PCCC across all three age groups at 0.71, 0.73, and 0.61 for adult, child, and neonatal age groups. GPT-3.5 had the next highest PCCC for the adult and child age groups at 0.62 and 0.69 PCCC respectively, performing slightly worse than GPT-4 by 0.09 and 0.04 PCCC. InSilicoVA had the second highest PCCC for the neonatal age group at 0.56, which was 0.05 PCCC worse than GPT-4. Both InterVA-5 and InSilicoVA had the lowest PCCC for the adult and child age groups at 0.41 and 0.45 PCCC respectively, while InterVA-5 and GPT-3.5 had the lowest PCCC for the neonatal age group at 0.49 and 0.51 respectively. The PCCC for models had larger differences between the minimum and maximum PCCC for the adult and child age groups at 0.3 and 0.28 respectively than the neonatal age group at 0.12.

@@

\clearpage
\section{Miscellaneous Performance}\label{app-perf-misc}

>>>>

\author*[1]{\fnm{Richard} \sur{Wen}}\email{richard.wen@utoronto.ca}

\author[1,2]{\fnm{Anteneh Tesfaye} \sur{Assalif}}\email{antenehta@gmail.com}

\author[1]{\fnm{Rajeev} \sur{Kamadod}}\email{rajeevk@kentropy.com}

@@

\author[1]{\fnm{Andy Sze-Heng} \sur{Lee}}\email{andylee@cs.toronto.edu}

@@

\author[1]{\fnm{Asha} \sur{Behdinan}}\email{asha.behdinan@mail.utoronto.ca}

@@

\author[1]{\fnm{Areeba} \sur{Zubair}}\email{areeba.zubair@mail.utoronto.ca}

\author[1]{\fnm{Thomas Kai Sze} \sur{Ng}}\email{kaisze.ng@unityhealth.to}

\author[1]{\fnm{Paijani} \sur{Sheth}}\email{paijani.sheth@mail.utoronto.ca}

\author[1]{\fnm{Patrick} \sur{Brown}}\email{patrick.brown@utoronto.ca}

\author[1]{\fnm{Prabhat} \sur{Jha}}\email{prabhat.jha@utoronto.ca}

\author[2]{\fnm{Rashid} \sur{Ansumana}}\email{rashidansumana@gmail.com}

\affil*[1]{\orgdiv{Centre for Global Health Research, St. Michael's Hospital}, \orgname{Unity Health Toronto and University of Toronto}, \orgaddress{\street{30 Bond St}, \city{Toronto}, \postcode{M5B 1W8}, \state{Ontario}, \country{Canada}}}

\affil[2]{\orgdiv{School of Community Health Sciences}, \orgname{Njala University}, \orgaddress{\city{Bo}, \country{Sierra Leone}}}

@@

The CSMF accuracy and PCCC metrics were calculated and compared for each model overall and by age group, followed by age ranges and COD for each age group as model performance can vary across ages and specific causes \cite{benaraEvaluationMethodsAssigning2023,murrayUsingVerbalAutopsy2014,setelValidityVerbalAutopsy2006}. Metrics were calculated overall for three age groups according to the CGHR-10 codes: adult (12 to 69 years), child (28 days to 11 years), and neonatal (under 28 days). For each of the adult and child age groups, metrics were calculated for five-year age ranges for records with ages at death of one-year or older and five-month age ranges for 28 days or older. For the neonatal age group, the age ranges of 0-6 days and 7-27 days were used. Metrics were also calculated by CODs defined by CGHR-10 codes, which include 19, 10, and 7 CODs for adult, child, and neonatal records respectively.

@@

Of the 6939 records, GPT-4 had the highest PCCC followed by GPT-3.5, InSilicoVA, and InterVA-5 at 0.71, 0.64, 0.44, and 0.43 PCCC respectively (Table \ref{tab-perf-overall}). GPT-3.5 and GPT-4 had improvements ranging from 0.20-0.28 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.07 PCCC. CSMF accuracies were above 0.7 for all models at 0.89, 0.86, 0.74, and 0.78 for GPT-4, GPT-3.5, InSilicoVA, and InterVA-5 respectively. Figure \ref{fig-perf-agegroup} shows the PCCC performance across age groups with reference to the CSMF Accuracy. GPT-4 had the highest PCCC across all three age groups at 0.71, 0.73, and 0.61 for adult, child, and neonatal age groups. GPT-3.5 had the next highest PCCC for the adult and child age groups at 0.62 and 0.69 PCCC respectively, performing slightly worse than GPT-4 by 0.09 and 0.04 PCCC. InSilicoVA had the second highest PCCC for the neonatal age group at 0.56, which was 0.05 PCCC worse than GPT-4. Both InterVA-5 and InSilicoVA had the lowest PCCC for the adult and child age groups at 0.41 and 0.45 PCCC respectively, while InterVA-5 and GPT-3.5 had the lowest PCCC for the neonatal age group at 0.49 and 0.51 respectively. The PCCC for models had larger differences between the minimum and maximum PCCC for the adult and child age groups at 0.3 and 0.28 respectively than the neonatal age group at 0.12. For reference, all records were also coded by GPT-3.5, GPT-4, InterVA-5, and InSilicoVA for comparison to physician agreed records. For physician agreed records, models had 0.05 to 0.1 higher PCCC values and 0.01 to 0.05 higher CSMF accuracy values (see Figure \ref{fig-perf-allvsagree} in Appendix \ref{app-perf-misc}).

@@

\clearpage
\section{Performance By Other Variables}\label{app-perf-misc}
